---
title: "MA_youtube_eda"
date: "`r Sys.Date()`"
output:
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(stringr)
```




```{r}
data <- read.csv("/Users/tarikajain/Desktop/Marketing/youtube_channel_videos_details.csv")
head(data)
str(data)
summary(data)

```
#Seeing generally videos per year for each channel
```{r}
data$year <- year(data$published_at)

# Group by channel_id, channel_name, and year, then count the number of videos
videos_per_year <- data %>%
  group_by(channel_id, channel_name, year) %>%
  summarize(num_videos = n(), .groups = 'drop')  # Ensure groups are dropped after summarization

# Iterate over each unique channel name
for (channel_name in unique(videos_per_year$channel_name)) {
  # Filter data for the current channel
  channel_data <- videos_per_year[videos_per_year$channel_name == channel_name, ]
  
  # Convert year to a factor for discrete bar plotting
  channel_data$year <- as.factor(channel_data$year)

  # Create plot
  plot <- ggplot(channel_data, aes(x = year, y = num_videos)) +
    geom_bar(stat = "identity", fill = "firebrick", color = "black") +
    geom_vline(xintercept = as.character(2022), linetype = "dashed", color = "blue") +
    labs(title = paste("Number of Videos per Year for Channel", channel_name),
         x = "Year",
         y = "Number of Videos") +
    theme_minimal()
  
  # Print the plot
  print(plot)
}
```

Our data cover a long range of time. 


# Subsetting the data for 3 months before and after the launch of ChatGPT

```{r}
# Calculate the start and end date for the 3-month range
start_date <- as.Date("2022-08-30")  # Three months before November 30, 2022
end_date <- as.Date("2023-02-28")    # Three months after November 30, 2022

# Filter the data for the 6-month period
subset_data <- data %>%
  filter(published_at >= start_date & published_at <= end_date)

# View the subsetted data
head(subset_data)
str(subset_data)
subset_data <- subset_data %>%
  mutate(channel_id = as.factor(channel_id),
         channel_name = as.factor(channel_name))
summary(subset_data)
subset_data <- subset_data %>%
  filter(!channel_name %in% c("3Blue1Brown", "Siraj Raval"))
 # Dropping this channel as it has very few videos during this period. Also dropping siraj raval as it does not have videos 3 months before chatGPT
```
# Comparing number of videos before and after the launch of ChatGPT

```{r}
# Count the number of videos 2 months before and after November 30, 2022, for each channel
video_count_2_months <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    videos_before_2_months = sum(published_at >= as.Date("2022-09-30") & published_at < as.Date("2022-11-30")),
    videos_after_2_months = sum(published_at > as.Date("2022-11-30") & published_at <= as.Date("2023-01-30"))
  )

print(video_count_2_months)
```
This does not seem very insightful at this stage. 

```{r}

# Calculate the video counts
video_count_grouped <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    Videos_Before = sum(published_at < as.Date("2022-11-30")),
    Videos_After = sum(published_at >= as.Date("2022-11-30")),
    .groups = 'drop'
  )

# Reshape data from wide to long format for plotting
video_count_long <- video_count_grouped %>%
  pivot_longer(
    cols = c(Videos_Before, Videos_After),
    names_to = "Period",
    values_to = "Number_of_Videos"
  ) %>%
  mutate(Period = factor(Period, levels = c("Videos_Before", "Videos_After")))  # Set factor levels to control bar order

# Create the grouped bar chart
ggplot(video_count_long, aes(x = channel_name, y = Number_of_Videos, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(
    values = c("Videos_Before" = "#f9d448", "Videos_After" = "#7ca655"),
    labels = c("Videos_Before" = "Before ChatGPT", "Videos_After" = "After ChatGPT")
  )  +  # Custom colors; adjust as needed
  labs(
    title = "Number of Videos Before and After November 30, 2022 by Channel",
    x = "Channel Name",
    y = "Number of Videos",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))





```


```{r}
# Average view counts 3 months before and after November 30, 2022, for each channel
average_view_count_3_month <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    avg_views_before = mean(view_count[published_at < as.Date("2022-11-30")]),
    avg_views_after = mean(view_count[published_at >= as.Date("2022-11-30")])
  )

# Print the comparison
print(average_view_count_3_month)

# Average view counts 1 month before and after November 30, 2022, for each channel
average_view_count_1_month <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    avg_views_before_1_month = mean(view_count[published_at >= as.Date("2022-10-30") & published_at < as.Date("2022-11-30")]),
    avg_views_after_1_month = mean(view_count[published_at > as.Date("2022-11-30") & published_at <= as.Date("2022-12-30")])
  )

# Print the comparisons
print(average_view_count_1_month)

# Average view counts 2 months before and after November 30, 2022, for each channel
average_view_count_2_months <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    avg_views_before_2_months = mean(view_count[published_at >= as.Date("2022-09-30") & published_at < as.Date("2022-11-30")]),
    avg_views_after_2_months = mean(view_count[published_at > as.Date("2022-11-30") & published_at <= as.Date("2023-01-30")])
  )

# Print the comparisons
print(average_view_count_2_months)

#average view counts for the 3-month data
average_view_count_long_3_month <- average_view_count_3_month %>%
  pivot_longer(cols = starts_with("avg_views"), names_to = "Period", values_to = "Average Views") %>%
  mutate(Period = factor(Period, levels = c("avg_views_before", "avg_views_after"), labels = c("Before", "After")))

#  total views for before and after Nov 30, 2022, for each channel
total_views <- average_view_count_long_3_month %>%
  group_by(channel_name) %>%
  summarize(total_views = sum(`Average Views`))

# Join the totals back to the original data
average_view_count_long_3_month <- average_view_count_long_3_month %>%
  left_join(total_views, by = "channel_name")

# percentage of total views
average_view_count_long_3_month <- average_view_count_long_3_month %>%
  mutate(percentage = `Average Views` / total_views * 100)

# plot
ggplot(average_view_count_long_3_month, aes(x = channel_name, y = percentage, fill = Period)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Average View Counts 3 Months Before and After Nov 30, 2022",
       x = "Channel",
       y = "Percentage of Average View Count",
       fill = "Period") +
  scale_fill_manual(values = c("Before" = "#f9d448", "After" = "#7ca655"),
                    labels = c("videos_before" = "% Before ChatGPT", "videos_after" = "% After ChatGPT")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

```{r}
# Average like counts
average_like_count_3_months <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    avg_likes_before = mean(like_count[published_at < as.Date("2022-11-30")]),
    avg_likes_after = mean(like_count[published_at >= as.Date("2022-11-30")]),
    .groups = 'drop'
  )

# Average comment counts
average_comment_count_3_months <- subset_data %>%
  group_by(channel_name) %>%
  summarize(
    avg_comments_before = mean(comment_count[published_at < as.Date("2022-11-30")]),
    avg_comments_after = mean(comment_count[published_at >= as.Date("2022-11-30")]),
    .groups = 'drop'
  )


# percentages
average_like_count_long_3_months <- average_like_count_3_months %>%
  pivot_longer(cols = starts_with("avg_likes"), names_to = "Period", values_to = "Average Likes") %>%
  mutate(Period = factor(Period, levels = c("avg_likes_before", "avg_likes_after"), labels = c("Before", "After")))

total_likes <- average_like_count_long_3_months %>%
  group_by(channel_name) %>%
  summarize(total_likes = sum(`Average Likes`), .groups = 'drop')

average_like_count_long_3_months <- average_like_count_long_3_months %>%
  left_join(total_likes, by = "channel_name") %>%
  mutate(percentage = `Average Likes` / total_likes * 100)

# Plot 
ggplot(average_like_count_long_3_months, aes(x = channel_name, y = percentage, fill = Period)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Percentage of Average Like Counts 3 Months Before and After Nov 30, 2022",
       x = "Channel",
       y = "Percentage of Average Like Count",
       fill = "Period") +
  scale_fill_manual(values = c("Before" = "#f9d448", "After" = "#7ca655"),
                    labels = c("videos_before" = "% Before ChatGPT", "videos_after" = "% After ChatGPT")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Pivot the comment count data to long format and calculate percentages
average_comment_count_long_3_months <- average_comment_count_3_months %>%
  pivot_longer(cols = starts_with("avg_comments"), names_to = "Period", values_to = "Average Comments") %>%
  mutate(Period = factor(Period, levels = c("avg_comments_before", "avg_comments_after"), labels = c("Before", "After")))

total_comments <- average_comment_count_long_3_months %>%
  group_by(channel_name) %>%
  summarize(total_comments = sum(`Average Comments`), .groups = 'drop')

average_comment_count_long_3_months <- average_comment_count_long_3_months %>%
  left_join(total_comments, by = "channel_name") %>%
  mutate(percentage = `Average Comments` / total_comments * 100)

# Plot 
ggplot(average_comment_count_long_3_months, aes(x = channel_name, y = percentage, fill = Period)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Average Comment Counts 3 Months Before and After Nov 30, 2022",
       x = "Channel",
       y = "Percentage of Average Comment Count",
       fill = "Period") +
  scale_fill_manual(values = c("Before" = "#f9d448", "After" = "#7ca655")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```



# Separating AI-themed videos 



```{r}
# Keywords
keywords <- c("artificial intelligence", "chatgpt", "AI ", "machine learning", "neural")

videosAI <- subset_data %>%
  filter(str_detect(title, regex(keywords[1], ignore_case = TRUE)) |
           str_detect(description, regex(keywords[1], ignore_case = TRUE)) |
           str_detect(title, regex(keywords[2], ignore_case = TRUE)) |
           str_detect(description, regex(keywords[2], ignore_case = TRUE)) |
           str_detect(title, regex(keywords[3], ignore_case = TRUE)) |
           str_detect(description, regex(keywords[3], ignore_case = TRUE)))

head(videosAI)

# non-AI
videosNonAI <- anti_join(subset_data, videosAI)

head(videosNonAI)
```
### Now let's compare the video counts for AI related videos and non-AI related videos before and after the launch of ChatGPT

## AI video counts
```{r}

videosAI_count <- videosAI %>%
  group_by(channel_name) %>%
  summarize(
    Videos_Before_AI = sum(published_at < as.Date("2022-11-30")),
    Videos_After_AI = sum(published_at >= as.Date("2022-11-30")),
    .groups = 'drop'
  ) %>%
  pivot_longer(
    cols = c(Videos_Before_AI, Videos_After_AI),
    names_to = "Period",
    values_to = "Number_of_Videos"
  ) %>%
  mutate(Period = factor(Period, levels = c("Videos_Before_AI", "Videos_After_AI")))

# Plot
ggplot(videosAI_count, aes(x = channel_name, y = Number_of_Videos, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(
    values = c("Videos_Before_AI" = "#f9d448", "Videos_After_AI" = "#7ca655"),
    labels = c("Before ChatGPT", "After ChatGPT")
  ) +
  labs(
    title = "Number of AI-Related Videos Before and After November 30, 2022 by Channel",
    x = "Channel Name",
    y = "Number of Videos",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

# comparison at 1 month
AIvideo_count_1_month <- videosAI %>%
  group_by(channel_name) %>%
  summarize(
    videos_before_1_month = sum(published_at >= as.Date("2022-10-30") & published_at < as.Date("2022-11-30")),
    videos_after_1_month = sum(published_at > as.Date("2022-11-30") & published_at <= as.Date("2022-12-30"))
  )

print(AIvideo_count_1_month)

# comparison at 2 months
AIvideo_count_2_months <- videosAI %>%
  group_by(channel_name) %>%
  summarize(
    videos_before_2_months = sum(published_at >= as.Date("2022-09-30") & published_at < as.Date("2022-11-30")),
    videos_after_2_months = sum(published_at > as.Date("2022-11-30") & published_at <= as.Date("2023-01-30"))
  )


print(AIvideo_count_2_months)



```
There is clearly an increase in the number of videos being posted about AI for almost all of the channels. 

## NON-AI video counts

```{r}
videosNonAI_count <- videosNonAI %>%
  group_by(channel_name) %>%
  summarize(
    Videos_Before_NonAI = sum(published_at < as.Date("2022-11-30")),
    Videos_After_NonAI = sum(published_at >= as.Date("2022-11-30")),
    .groups = 'drop'
  ) %>%
  pivot_longer(
    cols = c(Videos_Before_NonAI, Videos_After_NonAI),
    names_to = "Period",
    values_to = "Number_of_Videos"
  ) %>%
  mutate(Period = factor(Period, levels = c("Videos_Before_NonAI", "Videos_After_NonAI")))

# Plot
ggplot(videosNonAI_count, aes(x = channel_name, y = Number_of_Videos, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(
    values = c("Videos_Before_NonAI" = "#f9d448", "Videos_After_NonAI" = "#7ca655"),
    labels = c("Before ChatGPT", "After ChatGPT")
  ) +
  labs(
    title = "Number of Non-AI Related Videos Before and After November 30, 2022 by Channel",
    x = "Channel Name",
    y = "Number of Videos",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

There is clearly a drop in the number of videos being posted about non-AI related content across each of the channels.

# Understanding engagement levels - average likes, comments and views per video


```{r}
videosAI$published_at <- as.Date(videosAI$published_at)
videosNonAI$published_at <- as.Date(videosNonAI$published_at)

comparison_date <- as.Date("2022-11-30")


calculate_averages <- function(df, value_column) {
  df %>%
    group_by(channel_name) %>%
    summarize(
      average_before = mean({{ value_column }}[published_at < comparison_date], na.rm = TRUE),
      average_after = mean({{ value_column }}[published_at >= comparison_date], na.rm = TRUE)
    )
}

# Averages
AI_avg_view <- calculate_averages(videosAI, view_count)
AI_avg_like <- calculate_averages(videosAI, like_count)
AI_avg_comment <- calculate_averages(videosAI, comment_count)

nonAI_avg_view <- calculate_averages(videosNonAI, view_count)
nonAI_avg_like <- calculate_averages(videosNonAI, like_count)
nonAI_avg_comment <- calculate_averages(videosNonAI, comment_count)

print(AI_avg_view)
print(AI_avg_like)
print(AI_avg_comment)

print(nonAI_avg_view)
print(nonAI_avg_like)
print(nonAI_avg_comment)



```


## Average likes

```{r}
# Calculating absolute average likes for AI and non-AI videos before and after the comparison date
AI_avg_like <- videosAI %>%
  group_by(channel_name) %>%
  summarize(
    average_like_before = mean(like_count[published_at < comparison_date], na.rm = TRUE),
    average_like_after = mean(like_count[published_at >= comparison_date], na.rm = TRUE),
    .groups = 'drop'
  )

NonAI_avg_like <- videosNonAI %>%
  group_by(channel_name) %>%
  summarize(
    average_like_before = mean(like_count[published_at < comparison_date], na.rm = TRUE),
    average_like_after = mean(like_count[published_at >= comparison_date], na.rm = TRUE),
    .groups = 'drop'
  )

avg_like_data <- rbind(
  data.frame(channel_name = AI_avg_like$channel_name, average_like_before = AI_avg_like$average_like_before, average_like_after = AI_avg_like$average_like_after, Type = "AI"),
  data.frame(channel_name = NonAI_avg_like$channel_name, average_like_before = NonAI_avg_like$average_like_before, average_like_after = NonAI_avg_like$average_like_after, Type = "Non-AI")
)

avg_like_data_long <- avg_like_data %>%
  pivot_longer(
    cols = starts_with("average_like_"),
    names_to = "Period",
    values_to = "Average Likes"
  )

avg_like_data_long$Period <- factor(avg_like_data_long$Period, levels = c("average_like_before", "average_like_after"), labels = c("Before ChatGPT", "After ChatGPT"))

# AI Plot
AI_data <- avg_like_data_long %>% 
  filter(Type == "AI")

ggplot(AI_data, aes(x = channel_name, y = `Average Likes`, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("Before ChatGPT" = "#f9d448", "After ChatGPT" = "#7ca655")) +
  labs(
    title = "Average Likes Per AI Video Before and After ChatGPT by Channel",
    x = "Channel Name",
    y = "Average Likes",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Non-AI Plot
NonAI_data <- avg_like_data_long %>% 
  filter(Type == "Non-AI")

ggplot(NonAI_data, aes(x = channel_name, y = `Average Likes`, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("Before ChatGPT" = "#f9d448", "After ChatGPT" = "#7ca655")) +
  labs(
    title = "Average Likes Per Non-AI Video Before and After ChatGPT by Channel",
    x = "Channel Name",
    y = "Average Likes",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Average comments

```{r}

# Calculateing the absolute average comments for AI and non-AI videos before and after the comparison date
AI_avg_comment <- videosAI %>%
  group_by(channel_name) %>%
  summarize(
    average_comment_before = mean(comment_count[published_at < comparison_date], na.rm = TRUE),
    average_comment_after = mean(comment_count[published_at >= comparison_date], na.rm = TRUE),
    .groups = 'drop'
  )

NonAI_avg_comment <- videosNonAI %>%
  group_by(channel_name) %>%
  summarize(
    average_comment_before = mean(comment_count[published_at < comparison_date], na.rm = TRUE),
    average_comment_after = mean(comment_count[published_at >= comparison_date], na.rm = TRUE),
    .groups = 'drop'
  )


avg_comment_data <- rbind(
  data.frame(channel_name = AI_avg_comment$channel_name, average_comment_before = AI_avg_comment$average_comment_before, average_comment_after = AI_avg_comment$average_comment_after, Type = "AI"),
  data.frame(channel_name = NonAI_avg_comment$channel_name, average_comment_before = NonAI_avg_comment$average_comment_before, average_comment_after = NonAI_avg_comment$average_comment_after, Type = "Non-AI")
)

avg_comment_data_long <- avg_comment_data %>%
  pivot_longer(
    cols = starts_with("average_comment_"),
    names_to = "Period",
    values_to = "Average Comments"
  )


avg_comment_data_long$Period <- factor(avg_comment_data_long$Period, levels = c("average_comment_before", "average_comment_after"), labels = c("Before ChatGPT", "After ChatGPT"))

# AI Plot
AI_data <- avg_comment_data_long %>% 
  filter(Type == "AI")

ggplot(AI_data, aes(x = channel_name, y = `Average Comments`, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("Before ChatGPT" = "#f9d448", "After ChatGPT" = "#7ca655")) +
  labs(
    title = "Average Comment Counts Per AI Video Before and After ChatGPT by Channel",
    x = "Channel Name",
    y = "Average Comment Counts",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Non-AI Plot
NonAI_data <- avg_comment_data_long %>% 
  filter(Type == "Non-AI")

ggplot(NonAI_data, aes(x = channel_name, y = `Average Comments`, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("Before ChatGPT" = "#f9d448", "After ChatGPT" = "#7ca655")) +
  labs(
    title = "Average Comment Counts Per Non-AI Video Before and After ChatGPT by Channel",
    x = "Channel Name",
    y = "Average Comment Counts",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
## average views

```{r}

# Calculating absolute average views for AI and non-AI videos before and after the comparison date
AI_avg_view <- videosAI %>%
  group_by(channel_name) %>%
  summarize(
    average_view_before = mean(view_count[published_at < comparison_date], na.rm = TRUE),
    average_view_after = mean(view_count[published_at >= comparison_date], na.rm = TRUE),
    .groups = 'drop'
  )

NonAI_avg_view <- videosNonAI %>%
  group_by(channel_name) %>%
  summarize(
    average_view_before = mean(view_count[published_at < comparison_date], na.rm = TRUE),
    average_view_after = mean(view_count[published_at >= comparison_date], na.rm = TRUE),
    .groups = 'drop'
  )


avg_view_data <- rbind(
  data.frame(channel_name = AI_avg_view$channel_name, average_view_before = AI_avg_view$average_view_before, average_view_after = AI_avg_view$average_view_after, Type = "AI"),
  data.frame(channel_name = NonAI_avg_view$channel_name, average_view_before = NonAI_avg_view$average_view_before, average_view_after = NonAI_avg_view$average_view_after, Type = "Non-AI")
)

avg_view_data_long <- avg_view_data %>%
  pivot_longer(
    cols = starts_with("average_view_"),
    names_to = "Period",
    values_to = "Average Views"
  )

avg_view_data_long$Period <- factor(avg_view_data_long$Period, levels = c("average_view_before", "average_view_after"), labels = c("Before ChatGPT", "After ChatGPT"))

# AI Plot for views
AI_view_data <- avg_view_data_long %>% 
  filter(Type == "AI")

ggplot(AI_view_data, aes(x = channel_name, y = `Average Views`, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("Before ChatGPT" = "#f9d448", "After ChatGPT" = "#7ca655")) +
  labs(
    title = "Average Views Per AI Video Before and After ChatGPT by Channel",
    x = "Channel Name",
    y = "Average Views",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Non-AI Plot for views
NonAI_view_data <- avg_view_data_long %>% 
  filter(Type == "Non-AI")

ggplot(NonAI_view_data, aes(x = channel_name, y = `Average Views`, fill = Period)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  scale_fill_manual(values = c("Before ChatGPT" = "#f9d448", "After ChatGPT" = "#7ca655")) +
  labs(
    title = "Average Views Per Non-AI Video Before and After ChatGPT by Channel",
    x = "Channel Name",
    y = "Average Views",
    fill = "Period"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```






